Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors with Application to Renal Imaging Data

### 2025-10-07 manuscript/enar/submission.tex
import fundata

\documentclass[12pt]{biom}

% Page setup
%\usepackage[margin=1in]{geometry}
%\usepackage{setspace}
%\doublespacing % ensures double spacing
%\setlength{\textheight}{9in} % ~25 lines per page with 12pt double spaced
%\setlength{\parskip}{0pt}
%\setlength{\parindent}{0.5in}

% Font
\renewcommand{\rmdefault}{ptm} % Times font (Biometrics-like)

% References (Biometrics style uses Chicago-like author-year)

\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\RequirePackage{graphicx}

%custom packages
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{xcolor}

% \usepackage{url} % not crucial - just used below for the URL 
% \usepackage{bibunits}
\usepackage{subcaption}
% \usepackage{adjustbox}
\usepackage{verbatim} % to comment out

%\theoremstyle{plain}
%\newtheorem{axiom}{Axiom}
%\newtheorem{claim}[axiom]{Claim}
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}%custom
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Assumption, Definition, Example,     %%
%% Notation, Property, Remark, Fact         %%
%% use \theoremstyle{definition}            %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\theoremstyle{definition}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem*{example}{Example}
%\newtheorem*{fact}{Fact}
%\newtheorem{remark}{Remark}

% Counters for figures/tables
%\usepackage{etoolbox}
%\newcounter{totalfigtab}
%\pretocmd{\begin{figure}}{\stepcounter{totalfigtab}}{}{}
%	\pretocmd{\begin{table}}{\stepcounter{totalfigtab}}{}{}

		% Title page

		\begin{document}

Biomedical studies often collect multi-modal data, as in the Emory University renal study \citep{changBayesianLatentClass2020, jangPrincipalComponentAnalysis2021}, which records multiple renogram curves (functional data) along with scalar side information for each kidney.
Such distinct yet related physiological signals can be represented as functional- and scalar-valued i.i.d. covariates in a linear regression model:
\begin{equation}	\label{eq:hybrid_regression_model}
	Y_i
	= 
	\boldsymbol{\beta}^\top \mathbf{Z}_i
	+ 
	\sum 
	\limits_{k=1}^K 
	\int  \beta_k(t) X_{ik}(t) dt + \epsilon_i,
	\quad
	i = 1, \ldots, n,
	%= \langle \boldsymbol{\alpha}, \mathbf{Z} \rangle +  \langle \beta, X \rangle_{\mathcal{F}}  + \epsilon
\end{equation}
where $Y_i$ is a scalar response, 
$X_{i1}(t), \ldots, X_{iK}(t) \in \mathbb{L}^2[0,1]$ are functional predictors, 
$\mathbf{Z}_i = (Z_{i1}, \ldots, Z_{ip})^\top$ is a Euclidean vector covariate,
and $\epsilon_i$ denotes observational noise.
%In other words, this is a scalar-on-hybrid regression where the hybrid covariates belong to $\bigl( \mathbb{L}^2[0,1] \bigr)^{  K} \times \mathbb{R}^p$.
For notational convenience, we assume throughout that the responses and predictors are centered, so the intercept term can be omitted.

In addition to the ill-posedness of the infinite-dimensional slope functions $\beta_k(t)$ and the high dimensionality of the predictors ($K + p$), a further challenge comes from the strong correlations between functional and scalar predictors. This paper addresses all three issues in a unified manner by introducing a hybrid partial least squares (PLS) regression framework defined on a novel hybrid Hilbert space.

	\subsection{Previous works and limitations}
	To address the ill-posedness stemming from the infinite-dimensionality of the functional components, common remedies include basis expansion,
	using power-series  \citep{goldsmithPenalizedFunctionalRegression2011},
	B-splines \citep{cardotSplineEstimatorsFunctional2003, caiPredictionFunctionalLinear2006}, or 
	wavelets \citep{zhaoWaveletbasedLASSOFunctional2012},
	and structure-aware regularization, such as roughness penalties

	For the high-dimensionality, a major solution is using derived inputs.
	A straightforward approach in this direction is principal component analysis (PCA) regression, which applies PCA separately to the functional- (by, for example, \citealt{happ_multivariate_2018}) and scalar predictors, and then performs a classical multivariate regression on the combined scores.
	PCA regression is well-studied for linear regression with functional predictors alone \citep{hallMethodologyConvergenceRates2007, reissFunctionalPrincipalComponent2007, febrero-bandeFunctionalPrincipalComponent2017}.
	However, since the PCA-derived inputs are not informed by the response variable,   they are not guaranteed to capture the core regression relationship with a small number of derived inputs.
	In terms of predictive power, partial least squares (PLS) regression is a powerful alternative. It iteratively constructs a set of orthogonal latent components from the predictors that have the maximal covariance with the response variable, and use the resulting scores for regression.
	%
	PLS for linear regression in the context of functional predictor alone was introduced by \citet{predaPLSRegressionStochastic2005} for the case of a single predictor. Motivated by regression on chemometric spectra, \citet{reissFunctionalPrincipalComponent2007}, \citet{aguileraUsingBasisExpansions2010}, and \citet{aguileraPenalizedVersionsFunctional2016} extended the framework by incorporating basis approximations and roughness penalties to promote smoothness. \citet{delaigleMethodologyTheoryPartial2012} provided the first thorough theoretical analysis, while \citet{saricamPartialLeastsquaresEstimation2022} proposed a computationally efficient procedure based on Golub-Kahan bidiagonalization. For a comprehensive review, see \citet{febrero-bandeFunctionalPrincipalComponent2017}. More recently, \citet{beyaztasRobustFunctionalPartial2022} extended the framework to accommodate multiple functional predictors.

	However,  these existing approaches overlook the potentially strong correlations between functional and scalar components, which can lead to multicollinearity and suboptimal predictive performance. Multimodal correlations have mostly been addressed within an unsupervised learning framework. For instance, \citet{kolar_graph_2014} studied the estimation of joint undirected graphical models for  functional and vector data, while \citet{gengJointNonparametricPrecision2020} considered joint precision matrix estimation for brain measurements and confounding scalar covariate. \citet{jangPrincipalComponentAnalysis2021} proposed a joint PCA method that accounts for correlations between functional and scalar data. However, the resulting components are still not informed by the response and may fail to capture  correlation with the outcome.

	\subsection{Our contributions}
	To address the gaps mentioned above, we unify the remedies proposed for these three issues: basis expansion, partial least squares, and accounting for correlations between the functional and scalar components.
	We propose a hybrid PLS regression framework that integrates functional and vector predictors in a principled and coherent manner. To extract predictive structure from these jointly observed and potentially correlated data types, we define a Hilbert space that treats the tuple of functional and vector components as a single hybrid object, equipped with a suitable inner product. The hybrid PLS direction is then obtained by iteratively maximizing the empirical covariance with the response, subject to a unit-norm constraint in this Hilbert space. Our framework is readily applicable to dense and irregular functional data and supports regularization techniques to prevent overfitting and reduce variance. We also provide the mathematical properties that justify our algorithm.

	\section{Background on partial least squares and its extension to hybrid predictors}

	For intuition, let us return to the high-dimensional Euclidean predictor setting $
	Y_i =  \boldsymbol{\beta}^\top \mathbf{Z}_i + \epsilon_i$.
	A common way to address ill-posedness and correlation is to approximate the high-dimensional vector $\mathbf{Z}_i$ using a low-dimensional vector $
	\bigr(
	\widehat{\rho}_1^{[1]}, \ldots,   \widehat{\rho}_1^{[L]}  
	\bigr)^\top \in \mathbb{R}^L.
	$ To retain the regression relationship, the $l$-th PLS direction $\hat{\boldsymbol{\xi}}_l$ solves:
	\begin{equation}\label{scalar_PLS_sample}
		\max_{ \mathbf{ h } } 
		\widehat{\operatorname{Cov}}^2
		\bigl(
		\{
		\langle 
		\mathbf{ h }, \mathbf{Z}_i
		\rangle
		, Y_i
		\}_{i=1}^n
		\bigr)~\text{s.t.}~\| \mathbf{ h } \|_2 = 1,~\mathbf{ h }^\top  \widehat{\operatorname{Cov}}^2
		\bigl(
		\{
		\mathbf{Z}_i
		\}_{i=1}^n
		\bigr) \hat{\boldsymbol{\xi}}_j = 0, \quad j = 1, \ldots, l-1,
	\end{equation}
	where the two $\widehat{\operatorname{Cov}}^2$ denote sample cross-covariance and sample covariance, respectively. A standard algorithm for solving this optimization problem, called nonlinear iterative partial least squares (NIPALS) is presented in Algorithm \ref{alg:scalar_pls}.
	\begin{algorithm} 
		\caption{Scalar partial least squares regression}\label{alg:scalar_pls}
		\begin{algorithmic}[1]
			\State Standardize each $\mathbf{Z}_1, \ldots, \mathbf{Z}_n$ so that each feature have mean zero and variance one. Standardize $Y_1, \ldots, Y_n$.
			\For{$l = 1, 2, \ldots, L$}
			\\
			\textbf{PLS direction and score estimation:}
			\State $\widehat{\boldsymbol{\xi}}^{[l]} \leftarrow
			\arg \max_{  \boldsymbol{\alpha}   } 
			\widehat{\operatorname{Cov}}^2
			\bigl(
			\{
			\langle 
			\boldsymbol{\alpha}, \mathbf{Z}^{[l]}_i
			\rangle
			, Y^{[l]}_i
			\}_{i=1}^n
			\bigr)~\text{s.t.}~\| \boldsymbol{\alpha} \|_2 = 1
			$
			\Comment{PLS direction}\label{alg:step:scalar_pls_direction}
			\State 
			$
			\widehat{\rho}_i^{[l]} \leftarrow \langle \hat{\boldsymbol{\xi}}^{[l]},   \mathbf{Z}^{[l]}_i \rangle, 
			~i=1,  \ldots, n
			$ \Comment{PLS score}
			\\
			\textbf{Residualization:}
			\State $\nu^{[l]} 
			\leftarrow
			\frac{
				\sum_{i=1}^n Y_i^{[l]} \widehat{\rho}_{i}^{[l]}}{
				\sum_{i=1}^n \widehat{\rho}_{i}^{[l]2}
			}$ \Comment{Least squares estimate} 
			%
			\State $ Y_i^{[l+1]} \leftarrow Y_i^{[l]} - \nu^{[l]}\widehat{\rho}_i^{[l]}~i=1,  \ldots, n$
			%
			\State $ \widehat{\boldsymbol{\delta}}^{[l]}  \leftarrow \frac{1}{\sum_{i=1}^n \widehat{\rho}_{i}^{[l]2}}\sum_{i=1}^n  \widehat{\rho}_{i}^{[l]}\mathbf{Z}_i^{[l]} $ \Comment{Least squares estimate} 
			\label{alg:step:scalar_pls_residualization}
			\State $\mathbf{Z}_i^{[l+1]}  \leftarrow \mathbf{Z}_i^{[l]} -   \widehat{\rho}_{i}^{[l]}  \widehat{\boldsymbol{\delta}}^{[l]} $
			\EndFor
			\textbf{Regression coefficient estimation:}
			\State
			$\widehat{\boldsymbol{\iota}}^{[1]} \leftarrow \widehat{\boldsymbol{\xi}}^{[1]}$
			\For{$l =  2, \ldots, L$}
			\State $\widehat{\boldsymbol{\iota}}^{[l]} \leftarrow \widehat{ \boldsymbol{\xi} }^{[l]} - \sum_{u=1}^{l-1} \langle \widehat{\boldsymbol{\delta}}^{[u]}, \widehat{\boldsymbol{\xi}}^{[l]} \rangle 
			\widehat{\boldsymbol{\iota}}^{[u]}$  
			\EndFor
			$	\widehat{\boldsymbol{\beta}} \leftarrow \sum \limits_{l=1}^L\widehat{\nu}^{[l]} \widehat{\boldsymbol{\iota}}^{[l]}$ \Comment{regression coefficient estimate}
			\State \textbf{Output:} the  regression coefficient estimate
		\end{algorithmic}
	\end{algorithm}

	To extend Algorithm \ref{alg:scalar_pls} to the hybrid regression model \eqref{eq:hybrid_regression_model}, we must compute hybrid objects in both the PLS direction estimation and predictor residualization steps (lines \ref{alg:step:scalar_pls_direction} and \ref{alg:step:scalar_pls_residualization}). These computations cannot be separated into functional and scalar parts, as their correlation must be addressed.  
	A naive approach is to add loops over the observed evaluation points $t$ to lines \ref{alg:step:scalar_pls_direction} and \ref{alg:step:scalar_pls_residualization}, treating  
	\[
	\bigl( X_{i1}(t), \ldots, X_{iK}(t), Z_{i1}, \ldots, Z_{ip} \bigr) \in \mathbb{R}^{K+P}
	\]  
	as a Euclidean predictor, applying steps \ref{alg:step:scalar_pls_direction} and \ref{alg:step:scalar_pls_residualization}, and aggregating the results. This pointwise method, however, is computationally prohibitive for densely observed functional predictors, infeasible for irregular data, and prone to variability across the domain, resulting in overfitting and unstable predictions.  
	The main challenge of extension to the hybrid setting lies in the following points. First, 
	independent variables consist of multiple highly structured images and scalar predictors. 
	Second, our sample size is small compared to the dimension and number of functional and scalar predictors.
	Third,  existing partial least squares (PLS) methods can only accommodate (i) univariate or multivariate functional predictors without any scalar predictors \citep{predaPLSRegressionStochastic2005, delaigleMethodologyTheoryPartial2012, febrero-bandeFunctionalPrincipalComponent2017, Beyaztas2020}; or (ii) a univariate functional predictor with other scalar predictors \citep{Wang2018}.
	The next section proposes a new extension framework that addresses these issues.

	\section{Proposed PLS Algorithm} \label{section:main:our_algorithm}
	Our new framework builds on the fact that all computations in Algorithm \ref{alg:scalar_pls} are arithmetic operations in an inner product space. We therefore formally define the Hilbert space of hybrid random predictors along with its corresponding complete inner product, and leverage them to extend Algorithm \ref{alg:scalar_pls}.
	\begin{definition}[Hybrid space]\label{def:hilbert_space}
		Let $\mathbb{H}$ be a product space defined as the Cartesian product of $K$ copies of the space of square-integrable functions on $[0,1]$ and the $p$-dimensional Euclidean space:
		$$
		\mathbb{H} := (\mathbb{L}^2[0,1])^K \times \mathbb{R}^p.
		$$
		An element $h \in \mathbb{H}$ is an ordered tuple $h = (f_1, \dots, f_K, \mathbf{u})$, where $f_k \in \mathbb{L}^2[0,1]$ for $k=1, \dots, K$ and $\mathbf{u} \in \mathbb{R}^p$. This space is equipped with element-wise vector addition and scalar multiplication.
		We define an inner product on $\mathbb{H}$ for any two elements $h_1 = (f_1, \dots, f_K, \mathbf{u})$ and $h_2 = (g_1, \dots, g_K, \mathbf{v})$ as:
		\begin{equation}\label{eq: hybrid inner product}
			\langle h_1, h_2 \rangle_{\mathbb{H}} := \sum_{k=1}^K \int_0^1 f_k(t) g_k(t) \, dt + \omega \mathbf{u}^\top \mathbf{v},
		\end{equation}
		where $\omega$ is a positive constant ($\omega > 0$).
		The inner product induces a norm $\Vert \cdot \Vert_{\mathbb{H}}$ on the space, defined as $\Vert h \Vert_{\mathbb{H}} := \langle h, h \rangle_{\mathbb{H}}^{1/2}$, and a corresponding metric $d(h_1, h_2) = \Vert h_1 - h_2 \Vert_{\mathbb{H}}$.
	\end{definition}
	In \eqref{eq: hybrid inner product}, 
	$\omega$ is a positive weight that needs to be pre-specified or estimated. It is mainly used to take into account heterogeneity between functional and scalar parts in terms of measurement scale and/or amount of variation (see Section \ref{subsec: Data Preprocessing}).
	Without loss of generality and for the clarity of illustration, all the following theoretical results will be derived for $\omega=1$. The results remain valid for any positive weights. 
	\begin{lemma}[Hybrid Hilbert space]\label{lemma:hybrid_hilbert_space}
		The hybrid space $\mathbb{H}$ is a separable Hilbert space.
	\end{lemma} 
	Proof of Lemma \ref{lemma:hybrid_hilbert_space} is provided in Appendix \ref{section:proof:lemma:hybrid_hilbert_space}.

>>>>

\documentclass[referee,12pt,usenatbib]{biom}
\usepackage{booktabs}

\usepackage{bbm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{hyperref}
%\RequirePackage{,,}
\bibliographystyle{biom} % Biometrics bib style if installed
%\usepackage[authoryear,round]{natbib}
\title{Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors with Application to Renal Imaging Data}

\begin{document}
	\begin{abstract}
Motivated by renal imaging data with accompanying side information, we propose a hybrid partial least squares (PLS) regression method to capture cross-modality correlations with both functional and scalar predictors. At the population level, we extend the nonlinear iterative PLS (NIPALS) algorithm to a hybrid Hilbert space, unifying functional and scalar predictors into a single vector and iteratively maximizing the empirical covariance. We establish orthogonality of the PLS scores and $L_2$ convergence of the PLS fitted value to the true response. We then propose a practical sample-level algorithm incorporating regularization to promote smoothness of regression coefficients and PLS directions. The resulting low-dimensional representations effectively capture within- and between-modality variation and predictor-response correlation, while preserving key PLS properties such as orthonormality of PLS directions and orthogonality of PLS scores. Moreover, the procedure is computationally efficient, requiring only the solution of linear systems at each iteration.
\end{abstract}

\begin{keywords}
Dimension reduction;
Functional data analysis;
Multiple data modalities;
Multivariate data analysis;
Multivariate functional data;
Partial least square
\end{keywords}

\maketitle
\newpage
\section{Introduction} \label{sec: Introduction}
Biomedical studies often collect multi-modal data.
Our running example is the Emory University renal study \citep{changBayesianLatentClass2020, jangPrincipalComponentAnalysis2021}, which records two renogram curves (functional data) along with  side information (Euclidean vector) for each kidney.
Such distinct yet related physiological signals can be represented as functional- and scalar-valued i.i.d. covariates in a linear regression model:
\begin{equation}	\label{eq:hybrid_regression_model}
	Y_i
	= 
	\boldsymbol{\beta}^\top \mathbf{Z}_i
	+ 
	\sum 
	\limits_{k=1}^K 
	\int  \beta_k(t) X_{ik}(t) dt + \epsilon_i,
	\quad
	i = 1, \ldots, n,
	%= \langle \boldsymbol{\alpha}, \mathbf{Z} \rangle +  \langle \beta, X \rangle_{\mathcal{F}}  + \epsilon
\end{equation}
where $Y_i$ is a scalar response, 
$X_{i1}(t), \ldots, X_{iK}(t) \in \mathbb{L}^2[0,1]$ are functional predictors, 
$\mathbf{Z}_i = (Z_{i1}, \ldots, Z_{ip})^\top$ is a Euclidean vector covariate,
and $\epsilon_i$ denotes observational noise.
For notational convenience, we assume throughout that the responses and predictors are centered, so the intercept term can be omitted.
The key   challenge here is the strong correlations between functional and scalar predictors,
in addition to the ill-posedness of the infinite-dimensional slope functions $\beta_k(t)$ and the high dimensionality of the predictors ($K + p$). This paper addresses all three issues in a unified manner by introducing a hybrid partial least squares (PLS) regression framework defined on a novel hybrid Hilbert space.

	\paragraph{Previous works and limitations}
Basis expansion with regularization of roughness addresses the ill-posedness of functional data by exploiting their smoothness \citep{cardotSplineEstimatorsFunctional2003,aguileraPenalizedVersionsFunctional2016, 
	caiPredictionFunctionalLinear2006, zhaoWaveletbasedLASSOFunctional2012}.
High dimensionality can be addressed by using derived inputs, with PCA regression as a straightforward example. We can apply PCA  separately to functional (e.g., \citealt{happ_multivariate_2018}) and scalar predictors, followed by multivariate regression on the combined scores. While well studied for functional predictors alone \citep{hallMethodologyConvergenceRates2007, reissFunctionalPrincipalComponent2007, febrero-bandeFunctionalPrincipalComponent2017}, PCA regression may fail to capture the core regression relationship since the derived inputs are not guided by the response.
Partial least squares (PLS) regression is a powerful alternative that constructs orthogonal latent components from predictors to maximize covariance with the response.
For functional predictors, \citet{predaPLSRegressionStochastic2005} introduced PLS regression for a single predictor, with subsequent extensions incorporating
basis approximations and roughness penalties
\citep{reissFunctionalPrincipalComponent2007, aguileraUsingBasisExpansions2010, aguileraPenalizedVersionsFunctional2016}, and multiple functional predictors \citep{beyaztasRobustFunctionalPartial2022}.
Theoretical and computational advances were developed by \citet{delaigleMethodologyTheoryPartial2012} and \citet{saricamPartialLeastsquaresEstimation2022}, respectively; see \citet{febrero-bandeFunctionalPrincipalComponent2017} for a review.
These approaches ignore correlations between functional and scalar components, leading to multicollinearity and unreliable prediction. Multimodal correlations have mainly been explored in unsupervised settings,
such as graphical models, precision matrix estimation, and joint PCA \citep{kolar_graph_2014, gengJointNonparametricPrecision2020, jangPrincipalComponentAnalysis2021}. However,  these methods are response-agnostic and may miss outcome-relevant correlations.

\paragraph{Our contributions}
To address these gaps, we propose a hybrid PLS regression framework that unifies basis expansion, PLS, and functional-scalar correlations. Treating functional and Euclidean components as a single vector in a Hilbert space with a suitable inner product, we iteratively extract joint directions that maximize covariance with the response under a unit-norm constraint. The framework handles dense or irregular functional data, supports regularization to reduce overfitting, and comes with theoretical guarantees.

\section{Hybrid Hilbert Space and Population Version PLS Algorithm}
Our framework leverages that the nonlinear iterative PLS (NIPALS; \citealp{woldPathModelsLatent1975}) algorithm, a standard PLS fitting method, operates solely via inner-product space operations. We define a Hilbert space for hybrid predictors with a complete inner product to extend the algorithm.
%
\begin{theorem}[Hybrid space]\label{def:hilbert_space}
	An element $h \in \mathbb{H} := (\mathbb{L}^2[0,1])^K \times \mathbb{R}^p$ is an ordered tuple $h = (f_1, \dots, f_K, \mathbf{u})$, where $f_k \in \mathbb{L}^2[0,1]$ for $k=1, \dots, K$ and $\mathbf{u} \in \mathbb{R}^p$.  
We define an inner product on $\mathbb{H}$ for any two elements $h_1 = (f_1, \dots, f_K, \mathbf{u})$ and $h_2 = (g_1, \dots, g_K, \mathbf{v})$ as:
\begin{equation}\label{eq: hybrid inner product}
	\langle h_1, h_2 \rangle_{\mathbb{H}} := \sum_{k=1}^K \int_0^1 f_k(t) g_k(t) \, dt +   \mathbf{u}^\top \mathbf{v}.
\end{equation}
	The inner product induces a norm $\Vert \cdot \Vert_{\mathbb{H}}$ on the space, defined as $\Vert h \Vert_{\mathbb{H}} := \langle h, h \rangle_{\mathbb{H}}^{1/2}$, and a corresponding metric $d(h_1, h_2) = \Vert h_1 - h_2 \Vert_{\mathbb{H}}$.
	Then the hybrid inner product space $\mathbb{H}$ is a separable Hilbert space.
\end{theorem}

\begin{definition}[Hybrid Predictor]\label{def:hybrid_predictor}
	For the Hilbert space  $\mathbb{H}$ 
	defined in Definition \ref{def:hilbert_space},
	The Borel $\sigma$-field on $\mathbb{H}$, denoted $\mathfrak{B}(\mathbb{H})$, is the smallest $\sigma$-field containing the class $\mathfrak{M}$ of all sets of the form
	$
	\{q \in \mathbb{H} \mid \langle q, h \rangle \in O\},
	$
	for any $h \in \mathbb{H}$ and any open subset $O \subseteq \mathbb{R}$ (details can be found in Theorem 7.1.1 of \citealt{hsingTheoreticalFoundationsFunctional2015}).
	A hybrid predictor $ W_i = (X_{i1}(t), \ldots, X_{iK}(t), \mathbf{Z}_i)$ is a measurable mapping from a probability space $(\Omega, \mathfrak{F}, P)$ into   $
	\bigl(
	\mathbb{H}, \mathfrak{B}(\mathbb{H})
	\bigr)
	$. 
\end{definition}
Then the joint regression model \eqref{eq:hybrid_regression_model}  can be concisely written as
\begin{equation}		\label{eq: Hybrid functional model}
	Y_i = \langle \beta, W_i \rangle_{\mathbb{H}} + \epsilon_i,~\text{where}~\beta := \bigl( \beta_1(t), \ldots, \beta_K(t), \boldsymbol{\beta} \bigr)\in \mathbb{H}.
\end{equation}

We now derive the population version of NIPALS in the hybrid space.
At the population level, where the random objects are fully observable, we may drop the sample index $i$ and simply work with $Y$
and
$W = (X_1, \ldots, X_K, Z_1, \ldots, Z_p)$. Throughout this section we assume 
$\mathbb{E}[Y] = 0 \in \mathbb{R}$ and $\mathbb{E}[W] = 0 \in \mathbb{H}$.

The core optimization problem in PLS is formulated in terms of maximizing the squared cross-covariance. At the population level, this quantity is naturally characterized by cross-covariance operators, which we define next. %For notational simplicity, we suppress the iteration index $l$, with the understanding that for $l=1$ the predictors correspond to the original variables, while for $l \geq 2$ they represent the residualized versions.
\begin{definition}\label{def:cross_cov_terms}
	The covariance operator of the hybrid predictor $W$ is denoted as $\Sigma_{W} := \mathbb{E}[W \otimes W]$, so that for $u,v \in \mathbb{H}$, we have
	$
	\Sigma_{W} u = \mathbb{E}[
	\langle W, u\rangle_{\mathbb{H}}W
	],
	$
	$\langle \Sigma_{W} u, v \rangle_{\mathbb{H}}
	= \mathbb{E}[\langle W,u \rangle_{\mathbb{H}} \, \langle W,v \rangle_{\mathbb{H}}].
	$
	The cross-covariances between the response $Y$ and the predictors are
	\[
	\sigma_{YX}
	:=
	\bigl(\mathbb{E}[Y X_{1}],\ldots,\mathbb{E}[Y X_{K}]\bigr) \in \mathcal{F}^{K},
	\quad
	\sigma_{YZ} 
	:=
	\bigl(\mathbb{E}[Y Z_{1}],\ldots,\mathbb{E}[Y Z_{p}]\bigr)^\top \in \mathbb{R}^p.
	\]

	The cross-covariance between $Y$ and the hybrid predictor $W$ is then
	\[
	\Sigma_{YW} := \mathbb{E}[ Y W ] = \bigl( \sigma_{YX}, \, \sigma_{YZ} \bigr) \in \mathbb{H}.
	\]
\end{definition}
Now we explain the NIPALS algorithm at the population level.
Let $W^{[1]} := W$ and $Y^{[1]} := Y$.
For any   integer $l \geq 2$,
$W^{[l]}$ and $Y^{[l]}$ are residuals (see line 7,8 of Algorithm \ref{alg:population_pls}).
The PLS direction maximizes the squared covariance:
\begin{definition}\label{def:population_pls} Hybrid PLS direction at the $l$-th step,
	denoted $\xi^{[l]}$,
	is defined as
	\begin{equation}\label{const_eigen_pop}
		\xi^{[l]} := \arg\max_{h \in \mathbb{H}}
		\operatorname{Cov}^2(\langle W, h \rangle_{\mathbb{H}}, Y),~s.t.~\|h\|_{\mathbb{H}}=1, \langle h, \Sigma_W \, \xi^{[j]} \rangle_{\mathbb{H}} = 0,\, j < l.
	\end{equation}
\end{definition}
This problem can be solved by an iterative algorithm, and yields orthogonal scores:
\begin{theorem}[Properties of population PLS]
	\label{prop:residualization_equiv_eigen}
	The solution of optimization problem \eqref{const_eigen_pop} of Definition \ref{def:population_pls} is equivalent to that of population NIPALS (Algorithm \ref{alg:population_pls}).  For every \(s\ge 1\) and every \(j\le s-1\), we have
	$
	\mathbb{E}[\rho^{[j]}\rho^{[s]}]=0
	$
	and
	$
	\mathbb{E}[Y^{[s]}\rho^{[j]}]=0.
	$

\end{theorem}
\begin{algorithm} 
	\caption{Population NIPALS}\label{alg:population_pls}
	\begin{algorithmic}[1]
		\State
		$(W^{[1]}, Y^{[1]} \leftarrow (W, Y)$
		\For{$l = 1, 2, \ldots, L$}
		\State
		$
		\xi^{[l]}
		\leftarrow
		\arg\max_{h \in \mathbb{H}} \operatorname{Cov}^2(\langle W^{[l]}, h \rangle_{\mathbb{H}}, Y^{[l]})~s.t.~ \|h\|_{\mathbb{H}} = 1,
		$\label{alg:line:pls_direction} \Comment{PLS direction}
		%%%%%%%%%
		\State 
		$
		\rho^{[l]} \leftarrow \langle \xi^{[l]}, W^{[l]} \rangle 
		$ \Comment{PLS score}
		%%%%%%%%%%
		\State
		$
		\delta^{[l]}
		\leftarrow
		\frac{1}{\mathbb{E}[ (\rho^{[l]})^2 ]}  \mathbb{E}[W^{[l]} \rho^{[l]}]
		$\label{population_delta}
		\Comment{Linear regression of 
			$W^{[l]}$
			on $\rho^{[l]}$}
		%%%%%%%%%%%
		\State
		$
		\nu^{[l]}
		\leftarrow
		\frac{1}{\mathbb{E}[ (\rho^{[l]})^2 ]}
		\mathbb{E}[Y^{[l]} \rho^{[l]}]
		$
		\Comment{
			Linear regression of 
			$Y^{[l]} $
			on $\rho^{[l]}$
		}
		%%%%%%%%%%%%%%%%%%%
		\State 
		$ W^{[l+1]} \leftarrow W^{[l]} - \rho^{[l]} \, \delta^{[l]}$
		\label{population_predictor_residualize}
		\Comment{
			Residualized predictor
		}
		\State
		$ Y^{[l]} \leftarrow Y^{[l-1]} - \nu^{[l-1]} \, \rho^{[l-1]}$
		\Comment{
			Residualized response
		}
		\EndFor
		\State \textbf{Output:}
		PLS directions $\xi^{[1]}, \ldots, \xi^{[L]}$
	\end{algorithmic}
\end{algorithm}
We show in Theorem \ref{theorem:population_PLS} that problem \eqref{const_eigen_pop} and Algorithm \ref{alg:population_pls} are well-defined under mild conditions by proving line \ref{alg:line:pls_direction} is well-defined. 
To this end, Lemmas \ref{lemma:cross_cov_functional} and \ref{lemma:composite_cross_cov}
define the relevant operators, omitting the iteration index for residualized predictors, responses, and operators.
\begin{lemma}\label{lemma:cross_cov_functional}
	Define the operator $
	\mathcal{C}_{YW}$ such that for   any $h=(f_1, \ldots, f_K, \mathbf{v}) \in \mathbb{H}$,
	\begin{equation*} \mathcal{C}_{YW} h := \mathbb{E} \left[ \langle W, h \rangle_\mathbb{H} Y \right] = \langle \Sigma_{YW}, h \rangle_\mathbb{H} \in \mathbb{R}.
	\end{equation*} 
	The operator $\mathcal{C}_{YW}$ is a compact operator
	if   there exist finite constants $Q_1$ and $Q_2$ such that
	\begin{align}\label{condition_for_compact}
		\max \limits_{k=1, \ldots, K} \sup \limits_{t \in [0,1]} \mathbb{E}
		[Y X_{k}](t)^2 < Q_1 \quad \textnormal{and} \quad  \max_{r=1, \ldots, p} \mathbb{E}
		[Y Z_{r}]^2 < Q_2.
	\end{align}
\end{lemma}

\begin{lemma}\label{lemma:composite_cross_cov}
	Define
	$\mathcal{U} := \mathcal{C}_{WY} \circ \mathcal{C}_{YW}: \mathbb{H} \rightarrow \mathbb{H}$ as an operator such that,
	for $h \in \mathbb{H}$,
	\begin{equation*}
		\mathcal{U} h 
		= 
		\mathcal{C}_{WY} ( \mathcal{C}_{YW} h) 
		=
		\mathcal{C}_{YW} (\langle \Sigma_{YW}, h \rangle_\mathbb{H}) 
		=  
		\Sigma_{YW} \, \langle \Sigma_{YW}, h \rangle_\mathbb{H} 
		= 
		(\Sigma_{YW} \otimes \Sigma_{YW}) h.
	\end{equation*}
	In other words, $\mathcal{U} = \Sigma_{YW} \otimes \Sigma_{YW}$.
	Then $\mathcal{U}$
	is a self-adjoint and positive-semidefinite operator.
	Under the conditions of Lemma \ref{lemma:cross_cov_functional}, $\mathcal{U}$ is a compact operator.
\end{lemma}
By the Hilbert-Schmidt theorem (e.g., Theorem 4.2.4 in \citealp{hsingTheoreticalFoundationsFunctional2015}), Lemma \ref{lemma:composite_cross_cov} guarantees the existence of a complete orthonormal system of eigenfunctions $\{\xi_{(u)}\}_{u \in \mathbb{N}}$ of $\mathcal{U}$ in $\mathbb{H}$ such that $\mathcal{U} \xi_{(u)} = \kappa_{(u)}  \xi_{(u)}$, where $\{\kappa_{(u)}\}_{u \in \mathbb{N}}$ are the corresponding sequence of eigenvalues that goes to zero as $u \rightarrow \infty$, that is, $\kappa_{(1)} \ge \kappa_{(2)} \ge \cdots \ge 0$.
The following theorem presents   ensures that line \ref{alg:line:pls_direction} of Algorithm \ref{alg:population_pls} is well-defined under mild conditions.
\begin{theorem}[Tucker's criterion]\label{theorem:population_PLS} Under the conditions of Lemma \ref{lemma:cross_cov_functional}, the constrained maximum
	$
		\max \limits_{\substack{ \Vert \xi \Vert_\mathbb{H}=1}} \operatorname{Cov}^2 \left(\langle W , \, \xi \rangle_\mathbb{H}, Y \right)
	$
	is attained by the eigenfunction associated with the largest eigenvalue of the operator $\mathcal{U}$.
\end{theorem}
Finally, the following theorem states that the fitted values from hybrid PLS converge to those of ordinary linear regression in the mean-squared sense. This shows that hybrid PLS, as a dimension-reduction technique, can effectively capture the predictive power of a full linear model when a sufficient number of components is used.
\begin{theorem}\label{thm:hybrid_pls_convergence}
	Let $Y_{\mathrm{PLS}, L} = \sum_{l=1}^L \nu^{[l]} \rho^{[l]}$ be the Hybrid PLS fitted value with $L$ components. Then, the Hybrid PLS fitted value converges to $Y$ in the mean-squared sense:
	\begin{equation*}
		\lim_{L \to \infty} \mathbb{E}\left[ \| Y_{\mathrm{PLS},L } - Y  \|_2^2 \right] = 0.
	\end{equation*}
\end{theorem}
